---
title: "Functions_warmup"
author: "Charles Hendrickson"
date: "8/12/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(palmerpenguins)
```

## 

```{r}

quarter_splits <- c(1.0, 1.1, 1.2, 1.1, 1.4, 1.5, 1.6, 1.4)

out <- vector(mode = "numeric", 
              length = (length(quarter_splits) -1))

for (i in 1:(length(quarter_splits) -1)) {
  q <- quarter_splits[i] + quarter_splits[i +1]
  out[i] <- q
}

out

```


# Another for loop example:
```{r}
birddog_sum <- function(bird, dog) {
  add_birddog <- bird + dog
  return(add_birddog)
}

birddog_sum(bird = 100, dog = 5)

# How to break this? - add a non-numeric value for bird or dog

birddog_sum(bird = 25, dog = 10)
```

```{r}
say_hello <- function(friend, weekday) {
  print(paste("Hello,", friend, "have a great", weekday))
}

say_hello(friend = "Jesse", weekday = "Thursday")
```


```{r}
find_multiple <- function(a, b) {
  my_multiplication <- a * b
  print(my_multiplication)
}

find_multiple(a = 10, b = 6)

find_multiple(a = 10, b = 6) + 12

```

### Functions with conditions 

```{r}
animal_age <- function(animal, age) {
  
  if (!animal %in% c("dog", "goat")) {
    warning("Animal must be a dog or a goat.")
  }
  
  if (animal == "dog") {
    creature_age <- age * 7
  }
  
  else if (animal == "goat") {
    creature_age <- age * 4.7
  }
  
  return(creature_age)
  
}

# animal_age(animal = "urchin", age = 10)


```

# Right way to look for rows with "in operator" or "%in%"
# For any row does the row match Adelie or Gentoo?
```{r}
adelie_gentoo <- penguins %>%
  filter(species %in% c("Adelie", "Gentoo"))

```

# WRONG way to look for rows
# Looking for Adelie in the first row and Gentoo in the second row... then Adelie in the third row. 
```{r}
adelie_gentoo_WRONG <- penguins %>%
  filter(species == c("Adelie", "Gentoo"))

```

## Combine functions & for loops 

```{r}
df_means <- function(df) {
  for (i in 1:ncol(df)) {
    col_mean <- mean(df[[i]], na.rm = TRUE)
    column_name <- colnames(df[i])
    print(paste("The mean of", column_name, "is", col_mean))
  }
}

df_means(mtcars)

# colnames(df[i]) will return that column name as a string

# Update this code to include the column name associated with the mean value from the column


```

```{r}
# Make a subset of only structural size variables from the penguins for Chinstrap

chinstrap_sizes <- penguins %>% 
  filter(species == "Chinstrap") %>% 
  select(bill_length_mm, bill_depth_mm, flipper_length_mm,
         body_mass_g)

df_means(df = chinstrap_sizes)


```


## Condition to check the class of an imput
```{r}
lunchtime <- function(tacos, price_per) {
  
  if (is.numeric(tacos) == FALSE) {
    stop("You must input a number of tacos.")
  }
  
  
  total_price <- tacos * price_per
  return(total_price)
}
  
lunchtime(tacos = 10, price_per = 1.25)



```


## Example: use values from a data frame to calculate something based on user input.

```{r}
# Calculate the groundwater transport rate.

gw_rate <- function(site) {
  
  gw_slope_depth <- data.frame(sitename = c("mountain", "prairie", "desert", "beach"),
                               depth = c(32, 41, 63, 2),
                               slope = c(11.2, 0.4, 0.8, 2.6))
  
  site_select <- dplyr::filter(gw_slope_depth, sitename == site)
# You can do the above as this too:
# gw_slope_depth %>% filter(sitename == site)
  
  transport_rate <- 1.4 * site_select$slope + 3.6 * site_select$depth
  
  return(transport_rate)
}

gw_rate(site = "prairie")



```

### Now using switch() 

```{r}
gw_rate_switch <- function(site) {
  
  gw_slope_depth <- switch(site,
                           "mountain" = c(32, 11.2),
                           "prairie" = c(41, 0.4),
                           "desert" = c(63, 0.8),
                           "beach" = c(2, 2.6))
  
  
  transport_rate <- 1.4 * gw_slope_depth[2] + 3.6 * gw_slope_depth[1]
  
  return(transport_rate)
}

gw_rate_switch(site = "beach")

```

### Logistic growth 

```{r}

logistic_growth <- function(K, N0, r, time) {
  Nt <- K / (1 +(K-N0) / N0 * exp(-r * time))
  print(Nt)
}

logistic_growth(K = 5000, N0 = 200, r = 0.42, time = 6)

time_vec <- seq(from = 0, to = 35, by = 0.1)
bacterial_growth <- logistic_growth(K = 5000, N0 = 200, r = 0.42, time = time_vec)


bacterial_df <- data.frame(time_vec, bacterial_growth)


ggplot(data = bacterial_df, aes(x= time_vec, y = bacterial_growth)) +
  geom_point(aes(color = time_vec))

```

```{r}
r_seq <- seq(from = 0.20, to = 0.40, by = 0.01)

time_vec <- seq(from = 0, to = 35, by = 0.1)

r_seq
time_vec

out_matrix <- matrix(ncol = length(r_seq), nrow = length(time_vec))

# N0 = 100, K = 6000

for (i in seq_along(r_seq)) {
  for (j in seq_along(time_vec)) {
    pop <- logistic_growth(K = 6000, N0 = 100, r = r_seq[i],
                           time_vec[j])
    out_matrix[j, i] <- pop
  }
}

# Wrangling

out_df <- data.frame(time = time_vec, out_matrix)

# Update column names

colnames(out_df) <- c("time", paste0("gr_", r_seq))

# make a long form (tidy version) 

out_df_long <- out_df %>% 
  pivot_longer(-time, names_to = "growth_rate", values_to = "population")

# Make in ggplot a LINE GRAPH geom_line for population (y-axis) over time(x-axis) with color depending on the growth rate.

ggplot(data = out_df_long, aes(x = time, y = population)) +
  geom_line(aes(color = growth_rate))




```































